Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-17 surfaces-write-up/surface-write-up.tex
changed title
editing ... sleepy ... very sleepywefoewifhoewifhw

\title{Canonical rings of $\bq$-divisors on projective space and Hirzebruch surfaces}

@@

\begin{rem}
\label{rem:proj-one-point-ind}
The method of proof for Theorem ~\ref{thm:proj-one-point}
can be immediately generalized to work inductively. If $D = D' + \alpha H$
for some hyperplane $\alpha H$, then we can use the proof above to show
the canonical ring $R_D$ is generated over $R_D'$ by the same
generators and with the same relations given in Theorem
~\ref{thm:proj-one-point}.
\end{rem}
\todo{Perhaps we should consider restating the above theorem in this form, and then say for simplicity we prove it when the base is 0?}
\begin{thm}
\label{thm:proj-effective}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \sum_{i = 1}^{n} \alpha_i D_i$, with $\alpha_i =
\frac{p_i}{q_i} \in \bq$ and each $D_i$ a hypersurface of
$\bp^m$, be an effective divisor on $\bp^m$.

\begin{proof}
We derive the result from the following inductive argument.
If $n = 0$, i.e.~ $D = 0$, then we are done.
Now, we look to inductively add hypersurfaces.
If $D' = \sum_{i = 0}^{n} \alpha_i C_i$ is some $\bq$-divisor
on $\bp^m$, then we look at $D = D' + C$ where $C$ is a
hypersurface of degree $\pdeg$. Now use the Veronese mapping
to view $D$ as a divisor on $\bp^{m ({m + \pdeg} \choose \pdeg - 1)}$
and the image of $C$ as the intersection of a Veronese variety
with a hyperplane. Then we can reduce inductively adding hypersurfaces
to the case of adding hyperplanes. However, this is accomplished
by the method used in Theorem ~\ref{thm:proj-one-point} in an
inductive manner as mentioned in Remark ~\ref{rem:proj-one-point-ind}.
\end{proof}

@@

\section{Conclusion}
This paper extends the work of O'Dorney on $\bp^1$ to $\bp^m$ and Hirzebruch Surfaces, bounding the degrees of generation and relations of canonical rings of arbitrary divisors $D$.  The methods for deducing bounds on generators and relations of $\hirz_k$ applied similar ideas to O'Dorney's method on $\bp^1$.

>>>>

\title{Canonical rings of $\bq$-divisors on Minimal Rational Surfaces}

@@

\begin{rem}
\label{rem:proj-one-point-ind}
The method of proof for Theorem ~\ref{thm:proj-one-point}
can be immediately generalized to work inductively. If $D = D' + \alpha H$
for some hyperplane $H$, then the proof of Theorem ~\ref{thm:proj-one-point}, mutatis mutandis, extends to show
the canonical ring $R_D$ is generated over $R_D'$ by the same
generators and with the same relations given in Theorem
~\ref{thm:proj-one-point}.
\end{rem}
\todo{Perhaps we should consider restating the above theorem in this form, and then say for simplicity we prove it when the base is 0?}
\begin{thm}
\label{thm:proj-effective}
Let $[x_0, \ldots, x_m]$ be a projective coordinate system for
$\bp^m$. Let $D = \sum_{i = 1}^{n} \alpha_i D_i \in \bq \otimes \bp^m$, with $\alpha_i =
\frac{p_i}{q_i} \in \bq$.

Then the canonical ring
$R_D$
is generated in degrees up to $\max_{1 \leq i \leq n}{q_i}$ with
relations generated in degrees up to $2 \max_{1 \leq i \leq n}{q_i}$.
\end{thm}

\begin{proof}
We derive the result from the following inductive argument.
If $n = 0$, i.e.~ $D = 0$, then we are done.
Now, we look to inductively add hypersurfaces.
If $D' = \sum_{i = 0}^{n} \alpha_i D_i$ is some $\bq$-divisor
on $\bp^m$, then we look at $D = D' + C$ where $C$ is a
hypersurface of degree $\pdeg$. \todo{Aaron: rewrite this to make sense} Now use the Veronese mapping
to view $C$ as a hypersurface on $\bp^{\binom{{m + \pdeg}}{ \pdeg} -1 }$
and the image of $C$ as the intersection of a Veronese variety
with a hyperplane. Then we can reduce inductively adding hypersurfaces
to the case of adding hyperplanes. However, this is accomplished
by the method used in Theorem ~\ref{thm:proj-one-point} in an
inductive manner as mentioned in Remark ~\ref{rem:proj-one-point-ind}.
\end{proof}

@@

\section{Conclusion}
This paper extends the work of O'Dorney \cite{dorney:canonical} on $\bp^1$ to $\bp^m$ and Hirzebruch Surfaces, bounding the degrees of generation and relations of canonical rings of arbitrary divisors $D$.  The methods for deducing bounds on generators and relations of $\hirz_k$ applied similar methods to that of O'Dorney's for $\bp^1$.
